---
title: Package momcalc
output:
  github_document:
    toc: true
    toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
library(momcalc)
```

# Introduction

Package `momcalc` includes different functions to *calculate moments* of some distributions. It is possible to calculate the moments of a normal, lognormal or gamma distribution symbolically. These distributions may be multivariate. The distribution or moments of the [BEGG distribution][BEGG] can be calculated numerically.
The package also provides a test concerning the modality of a distribution. It tests, if a one dimensional distribution with compact support can be unimodal and is based on the moments of that distribution.

You can install `momcalc` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("CharlotteJana/momcalc")
```

# Symbolical calculation of moments

# The BEGG distribution

The *bimodal extension of the generalized Gamma-Distribution* (BEGG) was first introduced in

* Bulut, Y. M., & Arslan, O. (2015). [A bimodal extension of the generalized gamma distribution][BEGG]. *Revista Colombiana de Estadística*, 38(2), 371-384.

It is a scale mixture of the generalized gamma distribution that is almost always bimodal. The two modes can have different shapes, depending on the parameters $α$, $β$, $δ_0$, $δ_1$, $η$, $ε$, $μ$ and $σ$.

The density function can be calculated with `dBEGG` and is given by
$$ 
f(x) = \begin{cases} 
      \frac{αβ}{2η^\frac{δ₁+1}{α}(1+ε)^{δ₁}Γ(\frac{δ₁+1}{αβ})}~(-x)^{δ₁}~e^{-\frac{(-x)^{αβ}}{η^β(1+ε)^{αβ}}}~~~~ &if~x<0\\
      \frac{αβ}{2η^\frac{δ₀+1}{α}(1-ε)^{δ₀}Γ(\frac{δ₀+1}{αβ})}~x^{δ₀}~e^{-\frac{x^{αβ}}{η^β(1-ε)^{αβ}}}~~~~ &if~x≥0\\
\end{cases} 
$$

The k-th raw moment is given by
$$
E(X^k) = \frac{(-1)^kη^{k/α}(1+ε)^{k+1}}{2}~\frac{Γ(\frac{δ₁+k+1}{αβ})}{Γ(\frac{δ₁+1}{αβ})}~+~\frac{η^{k/α}(1-ε)^{k+1}}{2}~\frac{Γ(\frac{δ₀+k+1}{αβ})}{Γ(\frac{δ₀+1}{αβ})}
$$
and can be calculated with `mBEGG`.

```{r, echo = FALSE}
par(mfrow=c(2, 2))

# example 1
x <- seq(-2, 2, .01)
y <- dBEGG(x, alpha = 2, beta = 2, delta0 = 1, delta1 = 4, eta = 1, eps = 0)
plot(x, y, type = "l", ylab = "", main = "example 1", xlab = "α = 2, β = 2, δ₀ = 1, δ₁ = 4, η = 1, ε = 0")

# example 2
x <- seq(-2, 3, .01)
y <- dBEGG(x, alpha = 2, beta = 1, delta0 = 0, delta1 = 2, eta = 1, eps = -0.5)
plot(x, y, type = "l", ylab = "", main = "example 2", xlab = "α = 2, β = 1, δ₀ = 0, δ₁ = 2, η = 1, ε = -0.5")

# example 3
x <- seq(-2.5, 1.5, .01)
y <- dBEGG(x, alpha = 3, beta = 2, delta0 = 4, delta1 = 2, eta = 2, eps = 0.3)
plot(x, y, type = "l", ylab = "", main = "example 3", xlab = "α = 3, β = 2, δ₀ = 4, δ₁ = 2, η = 2, ε = 0.3")

# example 4
x <- seq(-4, 2, .01)
y <- dBEGG(x, alpha = 2, beta = 1, delta0 = 0, delta1 = 0, eta = 1, eps = 0.7)
plot(x, y, type = "l", ylab = "", main = "example 4", xlab = "α = 2, β = 1, δ₀ = 0, δ₁ = 0, η = 1, ε = 0.7")

```

# Test if a distribution is unimodal

Function `is.unimoal` checks if an (unknown) distribution with compact support can be unimodal. It uses several inequalities that were introduced in

* Teuscher, F., & Guiard, V. (1995). [Sharp inequalities between skewness and kurtosis for unimodal distributions][Teuscher]. *Statistics & probability letters*, 22(3), 257-260.
* Johnson, N. L., & Rogers, C. A. (1951). [The moment problem for unimodal distributions][Johnson]. *The Annals of Mathematical Statistics*, 433-439.
* Simpson, J. A., & Welch, B. L. (1960). [Table of the bounds of the probability integral when the first four moments are given][Simpson]. *Biometrika*, 47(3/4), 399-410.

Depending on the inequality, moments up to order 2 or 4 are required. A distribution that satisfies all inequalities that contain only moments up to order 2 is called **2-b-unimodal**. A distribution that satisfies all inequalities that contain only moments up to order 4 is called **4-b-unimodal**. It is possible that a multimodal distribution satisfies all inequalities and is therefore 2- and even 4-bimodal. But if at least one of the inequalities is not satisfied, the distribution cannot be unimodal. In this case, the test returns `not unimodal` as result.

Here are some examples using the BEGG distribution:

```{r}
# example 1 (bimodal)
example1 <- mBEGG(order = 1:4, alpha = 2, beta = 2, delta0 = 1, delta1 = 4, eta = 1, eps = 0)
is.unimodal(-2, 2, example1)

# example 2 (bimodal)
example2 <- mBEGG(order = 1:4, alpha = 2, beta = 1, delta0 = 0, delta1 = 2, eta = 1, eps = -0.5)
is.unimodal(-2, 3, example2)

# example 3 (bimodal)
example3 <- mBEGG(order = 1:4, alpha = 3, beta = 2, delta0 = 4, delta1 = 2, eta = 2, eps = 0.3)
is.unimodal(-2.5, 1.5, example3[1:2]) # test with moments of order 1 and 2
is.unimodal(-2.5, 1.5, example3) # test with moments of order 1 - 4

# example 4 (unimodal)
example4 <- mBEGG(order = 1:4, alpha = 2, beta = 1, delta0 = 0, delta1 = 0, eta = 1, eps = 0.7)
is.unimodal(-4, 2, example4)

```

# License

Package `momcalc` is free open source software licensed under the [GNU Public License](https://www.gnu.org/licenses/#GPL) (GPL 2.0 or above). The software is provided as is and comes WITHOUT WARRANTY.


[BEGG]: https://www.researchgate.net/publication/280136422_A_Bimodal_Extension_of_the_Generalized_Gamma_Distribution

[Teuscher]: https://www.sciencedirect.com/science/article/pii/016771529400074I

[Simpson]: https://www.jstor.org/stable/2333310?seq=1#page_scan_tab_contents

[Johnson]: https://www.jstor.org/stable/2236630?seq=1#page_scan_tab_contents